---
title: "BIOL 452B Appendix"
author: "Kat Chhen"
output: html_document
date: "2023-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


```{r eval = FALSE, message = FALSE}
#install packages
devtools::install_github("ctmm-initiative/ctmm", force = TRUE) #if package needs to be updated
devtools::install_github("jmcalabrese/corrMove", force = TRUE) #if package needs to be installed, there is no CRAN version available
```


```{r message = FALSE}
#load packages
#data, visualization
library(readr)
library(ggplot2)
library(khroma)          #colour blind friendly colour palette
library(dplyr)           #data wrangling
library(tidyr)           #data wrangling
library(tibble)          #data wrangling
library(lubridate)       #round_date() for corrMove
#analysis
library(ctmm)            #continuous-time movement models
library(lme4)            #pairwise sex test to see if differences are significant using glmer()
library(corrMove)        #correlative movement
```


# Data Preparation
```{r message = FALSE}
# Set working directory
setwd("C:/Users/achhen/OneDrive - UBC/Github/giant-anteater")

# Import dataset
ORIGINAL_DATA <- read_csv("data/Anteaters_NoOutliers (original data).csv", col_types = cols(timestamp = "c", class = "c", identity = "c", ID = "c", .default = "d"))
#subset to specific individuals
ANTEATER <- ORIGINAL_DATA[which(ORIGINAL_DATA$ID %in% c("Alexander", "Annie", "Anthony", "Beto", "Bumpus",
                                                        "Cate", "Christoffer","Elaine", "Hannah","Jackson",
                                                        "Jane","Kyle", "Larry 267", "Little Rick", "Luigi",
                                                        "Makao", "Margaret", "Maria", "Puji", "Reid",
                                                        "Rodolfo", "Sheron", "Thomas")),]

# Convert dataset to a telemetry object
DATA <- as.telemetry("data/Anteaters_NoOutliers (original data).csv")

# Import supplementary data containing biological information
ORIGINAL_META <- read_csv("data/Anteater_Results_Final.csv")
#subset to specific individuals
META_DATA <- ORIGINAL_META[c(1:3,8:10,12,14,17,19,20,22,23,25:29,33:35,37,38),]

#correct a mismatch entry for 'Larry 267' and 'Larry'
BIO_DATA <- mutate(select(ORIGINAL_META, 1:3), ID = if_else(condition = ID == 'Larry',
                                                       true = 'Larry 267',
                                                       false = ID))

# Import environment data
ENVIRO_DATA <- read_csv("data/anteater_annotated.csv")
names(ENVIRO_DATA)[18] <- "tree_cover"
names(ENVIRO_DATA)[19] <- "temperature"

#data subsetting and carpentry
# Subset data to isolate individuals found only in specific sites
SITE_1 <- DATA[c(1,3,9,10,12,15,21,24,27,29,36,38)] #(n=12)
SITE_1_adult <- DATA[c(1,9,10,12,15,21,29,36,38)] #adult (n=9)
SITE_1_male <- DATA[c(1,3,12,21,24,27,38)] #male (n=7)
SITE_1_male_adult <- DATA[c(1,12,21,38)] #male adult (n=4)
SITE_1_female <- DATA[c(9,10,15,29,36)] #female adult (n=5)
SITE_2 <- DATA[c(2,8,19,22,25,28,30,31,37,41,42)] #(n=11)
SITE_2_adult <- DATA[c(2,8,19,22,25,28,30,31,41,42)] #adult (n=10)
SITE_2_male <- DATA[c(8,25,37,42,28)] #male (n=5)
SITE_2_male_adult <- DATA[c(8,25,42,28)] #male adult (n=4)
SITE_2_female <- DATA[c(2,22,30,19,31,41)] #female adult (n=6)
```


```{r include = FALSE}
#Load RDS files
#movement models
FIT_1 <- readRDS("RDS/movement model/FIT_1.RDS")
FIT_1_male <- readRDS("RDS/movement model/FIT_1_male.RDS")
FIT_1_male_adult <- readRDS("RDS/movement model/FIT_1_male_adult.RDS")
FIT_1_female <- readRDS("RDS/movement model/FIT_1_female.RDS")
FIT_2 <- readRDS("RDS/movement model/FIT_2.RDS")
FIT_2_male <- readRDS("RDS/movement model/FIT_2_male.RDS")
FIT_2_male_adult <- readRDS("RDS/movement model/FIT_2_male_adult.RDS")
FIT_2_female <- readRDS("RDS/movement model/FIT_2_female.RDS")
#AKDE overlap
AKDE_1 <- readRDS("RDS/movement model/AKDE_1.RDS")
AKDE_1_male <- readRDS("RDS/movement model/AKDE_1_male.RDS")
AKDE_1_male_adult <- readRDS("RDS/movement model/AKDE_1_male_adult.RDS")
AKDE_1_female <- readRDS("RDS/movement model/AKDE_1_female.RDS")
AKDE_2 <- readRDS("RDS/movement model/AKDE_2.RDS")
AKDE_2_male <- readRDS("RDS/movement model/AKDE_2_male.RDS")
AKDE_2_male_adult <- readRDS("RDS/movement model/AKDE_2_male_adult.RDS")
AKDE_2_female <- readRDS("RDS/movement model/AKDE_2_female.RDS")
#pairwise dataframe
DATA_pairwise_1 <- readRDS("RDS/pairwise/DATA_pairwise_1.RDS")
DATA_pairwise_1_adult <- readRDS("RDS/pairwise/DATA_pairwise_1_adult.RDS")
DATA_pairwise_2 <- readRDS("RDS/pairwise/DATA_pairwise_2.RDS")
DATA_pairwise_2_adult <- readRDS("RDS/pairwise/DATA_pairwise_2_adult.RDS")
DATA_pairwise <- readRDS("RDS/pairwise/DATA_pairwise.RDS")
DATA_pairwise_adult <- readRDS("RDS/pairwise/DATA_pairwise_adult.RDS")
#proximity analysis
DATA_proximity_1 <- readRDS("RDS/proximity/DATA_proximity_1.RDS")
DATA_proximity_2 <- readRDS("RDS/proximity/DATA_proximity_2.RDS")
DATA_proximity <- readRDS("RDS/proximity/DATA_proximity.RDS")
#distance
distance_pair1 <- readRDS("RDS/distance/distance_pair1.RDS")
distance_pair2 <- readRDS("RDS/distance/distance_pair2.RDS")
distance_pair3 <- readRDS("RDS/distance/distance_pair3.RDS")
distance_pair4 <- readRDS("RDS/distance/distance_pair4.RDS")
distance_pair5 <- readRDS("RDS/distance/distance_pair5.RDS")
distance_pair6 <- readRDS("RDS/distance/distance_pair6.RDS")
distance_pair7 <- readRDS("RDS/distance/distance_pair7.RDS")
distance_pair8 <- readRDS("RDS/distance/distance_pair8.RDS")
distance_pair9 <- readRDS("RDS/distance/distance_pair9.RDS")
distance_pair10 <- readRDS("RDS/distance/distance_pair10.RDS")
distance_pair11 <- readRDS("RDS/distance/distance_pair11.RDS")
distance_pair12 <- readRDS("RDS/distance/distance_pair12.RDS")
#correlative movement
cmAnteater_pair1 <- readRDS("RDS/correlative movement/cmAnteater_pair1.RDS")
cmAnteater_pair2 <- readRDS("RDS/correlative movement/cmAnteater_pair2.RDS")
cmAnteater_pair3 <- readRDS("RDS/correlative movement/cmAnteater_pair3.RDS")
cmAnteater_pair4 <- readRDS("RDS/correlative movement/cmAnteater_pair4.RDS")
cmAnteater_pair5 <- readRDS("RDS/correlative movement/cmAnteater_pair5.RDS")
cmAnteater_pair6 <- readRDS("RDS/correlative movement/cmAnteater_pair6.RDS")
cmAnteater_pair7 <- readRDS("RDS/correlative movement/cmAnteater_pair7.RDS")
cmAnteater_pair8 <- readRDS("RDS/correlative movement/cmAnteater_pair8.RDS")
cmAnteater_pair9 <- readRDS("RDS/correlative movement/cmAnteater_pair9.RDS")
cmAnteater_pair10 <- readRDS("RDS/correlative movement/cmAnteater_pair10.RDS")
cmAnteater_pair11 <- readRDS("RDS/correlative movement/cmAnteater_pair11.RDS")
cmAnteater_pair12 <- readRDS("RDS/correlative movement/cmAnteater_pair12.RDS")
```

# Data Analysis
## Home Range
```{r include = FALSE}
#Calculate the home range size of each giant anteater
```


```{r}
#subset data for home range related information
HOMERANGE <- META_DATA[,c(1:6,21:23)]

#Calculate home range size (HR is in mm, divide by 1 million to get km)
HOMERANGE$HR_km <- HOMERANGE$HR/1000000
HOMERANGE$HR_min_km <- HOMERANGE$HR_min/1000000
HOMERANGE$HR_max_km <- HOMERANGE$HR_max/1000000
mean_hr <- round(mean(HOMERANGE$HR_km), 2)
min_hr <- round(mean(HOMERANGE$HR_min_km), 2)
max_hr <- round(mean(HOMERANGE$HR_max_km), 2)

#male home range size
male_mean_hr <- round(mean(HOMERANGE$HR_km[HOMERANGE$Sex == "Male"]), 2)
male_min_hr <- round(mean(HOMERANGE$HR_min_km[HOMERANGE$Sex == "Male"]), 2)
male_max_hr <- round(mean(HOMERANGE$HR_max_km[HOMERANGE$Sex == "Male"]), 2)

#female home range size
female_mean_hr <- round(mean(HOMERANGE$HR_km[HOMERANGE$Sex == "Female"]), 2)
female_min_hr <- round(mean(HOMERANGE$HR_min_km[HOMERANGE$Sex == "Female"]), 2)
female_max_hr <- round(mean(HOMERANGE$HR_max_km[HOMERANGE$Sex == "Female"]), 2)
```

```{r}
ggplot(data = HOMERANGE) +
  geom_linerange(aes(xmin = HR_min_km, xmax = HR_max_km, y = ID,
                     linewidth = 1)) + 
  geom_vline(xintercept = mean_hr, linetype = "dotdash") +
  xlab(bquote('Home Range' (km^2))) +
  theme(panel.grid.major = element_blank(), #removes horizontal gridlines
        panel.grid.minor = element_blank(), #removes vertical gridlines
        legend.position="none") #removes legend
```

## Home Range Overlap

Calculate the home range overlap using the ```ctmm``` package.

### Movement Models
#### Site 1

```{r}
#fit movement models
# SITE 1 - MALE AND FEMALE (n=12)
GUESS_1 <- lapply(SITE_1[1:12], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_1 <- lapply(1:12, function(i) ctmm.select(SITE_1[[i]],GUESS_1[[i]]) )
names(FIT_1) <- names(SITE_1[1:12])
overlap(FIT_1)
#calculate AKDE overlap, create aligned UDs
AKDE_1 <- akde(SITE_1[1:12],FIT_1)
overlap(AKDE_1)

# SITE 1 - MALE (n=7)
GUESS_1_male <- lapply(SITE_1_male[1:7], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_1_male <- lapply(1:7, function(i) ctmm.select(SITE_1_male[[i]],GUESS_1_male[[i]]) )
names(FIT_1_male) <- names(SITE_1_male[1:7])
overlap(FIT_1_male)
#calculate AKDE overlap, create aligned UDs
AKDE_1_male <- akde(SITE_1[1:7],FIT_1_male)
overlap(AKDE_1_male)

# SITE 1 - MALE ADULT (n=4)
GUESS_1_male_adult <- lapply(SITE_1[1:4], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_1_male_adult <- lapply(1:4, function(i) ctmm.select(SITE_1_male_adult[[i]],GUESS_1_male_adult[[i]]) )
names(FIT_1_male_adult) <- names(SITE_1_male_adult[1:4])
overlap(FIT_1_male_adult)
#calculate AKDE overlap
AKDE_1_male_adult <- akde(SITE_1_male_adult[1:4],FIT_1_male_adult)
overlap(AKDE_1_male_adult)

# SITE 1 - FEMALE (n=5)
GUESS_1_female <- lapply(SITE_1_female[1:5], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_1_female <- lapply(1:5, function(i) ctmm.select(SITE_1_female[[i]],GUESS_1_female[[i]]) )
names(FIT_1_female) <- names(SITE_1_female[1:5])
overlap(FIT_1_female)
#calculate AKDE overlap, create aligned UDs
AKDE_1_female <- akde(SITE_1_female[1:5],FIT_1_female)
overlap(AKDE_1_female)

#RDS of site 1 movement model and AKDE overlap have been saved to "RDS/movement model" folder
```

#### Site 2

```{r}
# SITE 2 - MALE AND FEMALE (n=11)
GUESS_2 <- lapply(SITE_2[1:11], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_2 <- lapply(1:11, function(i) ctmm.select(SITE_2[[i]],GUESS_2[[i]]) )
names(FIT_2) <- names(SITE_2[1:11])
overlap(FIT_2)
#calculate AKDE overlap, create aligned UDs
AKDE_2 <- akde(SITE_2[1:11],FIT_2)
overlap(AKDE_2)

# SITE 2 - MALE (n=5)
GUESS_2_male <- lapply(SITE_2_male[1:5], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_2_male <- lapply(1:5, function(i) ctmm.select(SITE_2_male[[i]],GUESS_2_male[[i]]) )
names(FIT_2_male) <- names(SITE_2_male[1:5])
overlap(FIT_2_male)
#calculate AKDE overlap, create aligned UDs
AKDE_2_male <- akde(SITE_2[1:5],FIT_2_male)
overlap(AKDE_2_male)

# SITE 2 - MALE ADULT (n=4)
GUESS_2_male_adult <- lapply(SITE_2_male_adult[1:4], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_2_male_adult <- lapply(1:4, function(i) ctmm.select(SITE_2_male_adult[[i]],GUESS_2_male_adult[[i]]) )
names(FIT_2_male_adult) <- names(SITE_2_male_adult[1:4])
overlap(FIT_2_male_adult)
#calculate AKDE overlap, create aligned UDs
AKDE_2_male_adult <- akde(SITE_2_male_adult[1:4],FIT_2_male_adult)
overlap(AKDE_2_male_adult)

# SITE 2 - FEMALE (n=6)
GUESS_2_female <- lapply(SITE_2_female[1:6], function(b) ctmm.guess(b,interactive=FALSE) )
FIT_2_female <- lapply(1:6, function(i) ctmm.select(SITE_2_female[[i]],GUESS_2_female[[i]]) )
names(FIT_2_female) <- names(SITE_2_female[1:6])
overlap(FIT_2_female)
#calculate AKDE overlap, create aligned UDs
AKDE_2_female <- akde(SITE_2_female[1:6],FIT_2_female)
overlap(AKDE_2_female)

#RDS of site 2 movement model and AKDE overlap have been saved to "RDS/movement model" folder
```

```{r}
#plot home range overlap (coloured based on sex; red = female, blue = male)
COL_1 <- c("#004488", "#004488", "#A50026", "#A50026", "#004488", "#A50026", "#004488", "#004488", "#004488", "#A50026", "#A50026", "#004488") 
COL_2 <- c("#A50026", "#004488", "#A50026", "#A50026", "#004488", "#004488", "#A50026", "#A50026", "#004488", "#A50026", "#004488") 
par(mfrow = c(1,2))
plot(AKDE_1, col.DF = COL_1, col.level = COL_1, col.grid = NA, level = NA)
title("A)", adj = 0)
title("Site 1")
plot(AKDE_2, col.DF = COL_2, col.level = COL_2, col.grid = NA, level = NA) 
title("B)", adj = 0)
title("Site 2")
```


### Pairwise Sex Comparison of Home Range Overlaps

```{r include = FALSE}
#pair movement models
FIT_pair1 <- FIT_1[c(5,8)]
FIT_pair2 <- FIT_1[c(5,6)]
FIT_pair3 <- FIT_1[c(3,8)]
FIT_pair4 <- FIT_1[c(6,9)]
FIT_pair5 <- FIT_1[c(3,10)]
FIT_pair6 <- FIT_1[c(3,11)]
FIT_pair7 <- FIT_1[c(6,12)]
FIT_pair8 <- FIT_2[c(1,5)]
FIT_pair9 <- FIT_2[c(5,9)]
FIT_pair10 <- FIT_2[c(7,11)]
FIT_pair11 <- FIT_2[c(9,11)]
FIT_pair12 <- FIT_2[c(8,10)]
#pair AKDE overlap
AKDE_pair1 <- AKDE_1[c(5,8)]
AKDE_pair2 <- AKDE_1[c(5,6)]
AKDE_pair3 <- AKDE_1[c(3,8)]
AKDE_pair4 <- AKDE_1[c(6,9)]
AKDE_pair5 <- AKDE_1[c(3,10)]
AKDE_pair6 <- AKDE_1[c(3,11)]
AKDE_pair7 <- AKDE_1[c(6,12)]
AKDE_pair8 <- AKDE_2[c(1,5)]
AKDE_pair9 <- AKDE_2[c(5,9)]
AKDE_pair10 <- AKDE_2[c(7,11)]
AKDE_pair11 <- AKDE_2[c(9,11)]
AKDE_pair12 <- AKDE_2[c(8,10)]
#proximity(AKDE_pair1, FIT_pair1, GUESS = (error = FALSE))
```


```{r}
#data carpentry
#Create home range overlap data frame in pairwise format

#Site 1
# Create a data frame from the overlap values
#calculate 95% overlap
overlap_1 <- overlap(object = AKDE_1, level = 0.95) 
#extract matrix from array (the median layer of the cube (aka array))
overlap_1_matrix <- overlap_1$CI[ , , 2] 
#remove duplicate values of the matrix (upper triangle & diagonal portion of the matrix)
overlap_1_matrix[upper.tri(overlap_1_matrix, diag = TRUE)] <- NA
#check to see if the matrix extraction was correct and successful
overlap_1_matrix

# Convert overlap matrix into dataframe
pairwise_1 <- as.data.frame(overlap_1_matrix)

#data carpentry
pairwise_1$anteater_A <- rownames(pairwise_1) 
pairwise_1 <- pivot_longer(pairwise_1, cols = -anteater_A, names_to = 'anteater_B', values_to = 'overlap', values_drop_na = TRUE)
#add biometric data to the dataframe
pairwise_1 <- left_join(pairwise_1, rename(BIO_DATA, anteater_A = ID), by = "anteater_A")
pairwise_1 <- left_join(pairwise_1, rename(BIO_DATA, anteater_B = ID), by = "anteater_B", suffix = c(".A", ".B"))

#add column to indicate which sexes that are being compared
DATA_pairwise_1 <- mutate(pairwise_1, 
                          sex_comparison = case_when(paste(Sex.A, Sex.B) == "Male Male" ~ "male-male",
                                                     paste(Sex.A, Sex.B) == "Female Female" ~ "female-female",
                                                     paste(Sex.A, Sex.B) == "Male Female" ~ "male-female",
                                                     paste(Sex.A, Sex.B) == "Female Male" ~ "male-female"))
#add a site column
DATA_pairwise_1$site <- "1"

#subset dataframe exclude subadults from dataframe matrix
DATA_pairwise_1_adult <- DATA_pairwise_1[which(DATA_pairwise_1$Age.A != "Subadult"),]
DATA_pairwise_1_adult <- DATA_pairwise_1_adult[which(DATA_pairwise_1_adult$Age.B != "Subadult"),]
DATA_pairwise_1_adult$site <- "1"

#SITE 2
# Create a data frame from the overlap values
#calculate 95% overlap
overlap_2 <- overlap(object = AKDE_2, level = 0.95) 

#extract median layer from array (the median layer of the cube (aka array), in the form of a matrix)
overlap_2_matrix <- overlap_2$CI[ , , 2]

#remove duplicate values of the matrix (upper triangle & diagonal portion of the matrix)
overlap_2_matrix[upper.tri(overlap_2_matrix, diag = TRUE)] <- NA
#check to see if the matrix extraction was correct and successful
overlap_2_matrix

# Convert overlap matrix into dataframe
pairwise_2 <- as.data.frame(overlap_2_matrix)

#data carpentry
pairwise_2$anteater_A <- rownames(pairwise_2)
pairwise_2 <- pivot_longer(pairwise_2, cols = -anteater_A, names_to = 'anteater_B', values_to = 'overlap', values_drop_na = TRUE)
pairwise_2 <- left_join(pairwise_2, rename(BIO_DATA, anteater_A = ID), by = "anteater_A")
pairwise_2 <- left_join(pairwise_2, rename(BIO_DATA, anteater_B = ID), by = "anteater_B", suffix = c(".A", ".B"))

#add column to indicate which sexes that are being compared
DATA_pairwise_2 <- mutate(pairwise_2,
                          sex_comparison = case_when(paste(Sex.A, Sex.B) == "Male Male" ~ "male-male",
                                                     paste(Sex.A, Sex.B) == "Female Female" ~ "female-female", 
                                                     paste(Sex.A, Sex.B) == "Male Female" ~ "male-female",
                                                     paste(Sex.A, Sex.B) == "Female Male" ~ "male-female"))
#add a site column
DATA_pairwise_2$site <- "2"

#subset dataframe to exclude subadults from dataframe matrix
DATA_pairwise_2_adult <- DATA_pairwise_2[which(DATA_pairwise_2$Age.A != "Subadult"),]
DATA_pairwise_2_adult <- DATA_pairwise_2_adult[which(DATA_pairwise_2_adult$Age.B != "Subadult"),]
DATA_pairwise_2_adult$site <- "2"

#combine both sites
DATA_pairwise <- bind_rows(DATA_pairwise_1, DATA_pairwise_2)
DATA_pairwise_adult <- bind_rows(DATA_pairwise_1_adult, DATA_pairwise_2_adult)

#RDS of pairwise have been saved to "RDS/pairwise" folder
```

```{r}
ggplot(data = DATA_pairwise, mapping = aes(x = sex_comparison, y = overlap, fill = sex_comparison)) + 
  geom_boxplot() +
  ylab("Home range overlap") +
  xlab("Sex") +
  #  ggtitle("HR Overlap pairwise comparison of sex") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), #removes horizontal gridlines
        panel.grid.minor = element_blank(), #removes vertical gridlines
        legend.position="none") + #removes legend
  scale_fill_manual(values = c("#A50026", "#9970AB", "#004488"),
                    labels = c("Female - Female", "Male - Female", "Male - Male"),
                    name = "") +
  scale_y_continuous(limits = c(0,1))
```


#### Pairwise Analysis

Determine if sex is a significant factor in the occurrence home range overlap using the ```lme4``` package.

```{r}
#test for significance in sex, compare model with and without sex as a variable
test <- glmer(overlap ~ sex_comparison + (1|site), family = binomial, data = DATA_pairwise)
test2 <- glmer(overlap ~ 1 + (1|site), family = binomial, data = DATA_pairwise)
anova(test, test2)
test_results <- anova(test, test2)
test_pvalue <- round(test_results$`Pr(>Chisq)`[2], 2)

```

## Proximity

Calculate the proximity statistics using the ```ctmm``` package. Proximity refers to the nearness in space at each time-interval.\
refer to help("proximity")

### Site 1

```{r}
# create empty columns for where the result information will be added/filled into
DATA_pairwise_1$proximity_low <- NA
DATA_pairwise_1$proximity_est <- NA
DATA_pairwise_1$proximity_high <- NA

#Note: this will take a while (several days depending on the data)
# Using tryCatch
for(i in 1:nrow(DATA_pairwise_1)){
  ANIMAL_A <- as.character(DATA_pairwise_1[i, 'anteater_A']) # add as.character due to tibble format
  ANIMAL_B <- as.character(DATA_pairwise_1[i, 'anteater_B'])
  TRACKING_DATA.1 <- DATA[c(ANIMAL_A, ANIMAL_B)] # extract anteater by name, has extra layers .-. it doesnt work, that is why
  # line above is using as.character removes all the fluff because you just want the text string
  MODELS.1 <- list(FIT_1[ANIMAL_A][[1]], FIT_1[ANIMAL_B][[1]])
  PROXIMITY1 <- tryCatch(
    {
      PROXIMITY_1 <- proximity(data = TRACKING_DATA.1, CTMM = MODELS.1, GUESS=ctmm(error=FALSE))},
    error=function(err){
      PROXIMITY_1 <- c(NA,NA,NA)
      return(PROXIMITY_1)
    }
  )
  DATA_pairwise_1[i, c("proximity_low")] <- PROXIMITY_1[1]
  DATA_pairwise_1[i, c("proximity_est")] <- PROXIMITY_1[2]
  DATA_pairwise_1[i, c("proximity_high")] <- PROXIMITY_1[3]
  write.csv(DATA_pairwise_1, "data/DATA_proximity_1.csv", row.names = FALSE)
  Sys.sleep(0.5) #this staggers each output by 0.5 seconds, a way to see the loop happening in real time
}

```


### Site 2
```{r}
# Using tryCatch
for(i in 1:nrow(DATA_pairwise_2)){
  ANIMAL_A <- as.character(DATA_pairwise_2[i, 'anteater_A']) # add as.character due to tibble format
  ANIMAL_B <- as.character(DATA_pairwise_2[i, 'anteater_B'])
  TRACKING_DATA_2 <- DATA[c(ANIMAL_A, ANIMAL_B)] # extract anteater by name, has extra layers .-. it doesnt work, that is why
  # line above is using as.character removes all the fluff because you just want the text string
  MODELS_2 <- list(FIT_2[ANIMAL_A][[1]], FIT_2[ANIMAL_B][[1]])
  PROXIMITY_2 <- tryCatch(
    {
      #calculate the proximity statistic
      PROXIMITY_2 <- proximity(data = TRACKING_DATA_2, CTMM = MODELS_2, GUESS=ctmm(error=FALSE))},
    error=function(err){
      PROXIMITY_2 <- c(NA,NA,NA)
      return(PROXIMITY_2)
    }
  )
  DATA_pairwise_2[i, c("proximity_low")] <- PROXIMITY_2[1]
  DATA_pairwise_2[i, c("proximity_est")] <- PROXIMITY_2[2]
  DATA_pairwise_2[i, c("proximity_high")] <- PROXIMITY_2[3]
  write.csv(DATA_pairwise_2, "data/DATA_proximity_2.csv", row.names = FALSE)
    Sys.sleep(0.5) #this staggers each output by 0.5 seconds, a way to see the loop happening in real time
}


```

```{r}
#combine the proximity of the two sites into one dataframe
DATA_proximity <- bind_rows(DATA_proximity_1, DATA_proximity_2)

#identify the pairs that had a proximity above/below 1
#above 1 indicates further apart
#below 1 indicates closer together
proximity_above1 <- DATA_proximity[which(DATA_proximity$proximity_low > 1),]
proximity_below1 <- DATA_proximity[which(DATA_proximity$proximity_high < 1),]

#remove pairs with no overlap
proximity_below1 <- proximity_below1[proximity_below1$overlap !=0,]

#RDS of proximity have been saved to "RDS/proximity" folder
```

```{r}
ggplot(data = DATA_proximity, aes(y = proximity_est, x = overlap, col = sex_comparison)) +
  geom_hline(yintercept = 1, col = "grey70", linetype = "dashed") +
  geom_point(size = 2) +
  geom_segment(aes(x = overlap, xend = overlap, y = proximity_low, yend = proximity_high), linewidth = 0.5) +
  scale_y_log10(expand = c(0,0.1)) +
  scale_x_continuous(limits = c(0,1), expand = c(0,0.02)) +
  scale_color_manual(values = c("#A50026", "#9970AB", "#004488"),
                     labels = c("Female - Female", "Male - Female", "Male - Male"),
                     name = "") +
  ylab("Proximity ratio") +
  xlab("Home range overlap") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size=8, family = "sans", face = "bold"),
        axis.title.x = element_text(size=8, family = "sans", face = "bold"),
        axis.text.y = element_text(size=6, family = "sans"),
        axis.text.x  = element_text(size=6, family = "sans"),
        legend.text = element_text(size=6, family = "sans", face = "bold"),
        plot.title = element_text(hjust = -0.05, size = 12, family = "sans", face = "bold"),
        legend.position = c(0.85, 0.15),
        legend.key=element_blank(),
        panel.background = element_rect(fill = "transparent"),
        legend.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.margin = unit(c(0.2,0.1,0.2,0.2), "cm"))
```

```{r include = FALSE}
### sex comparison
#sex comparison
#gamma distribution
```

### Distance

Calculated the instantaneous Euclidean distance between the individuals in the dyad using the ```ctmm``` package.\
Use telemetry data

```{r}
#subset pairs that had a proximity above/below 1
# 'DATA' needs to contain the pair ie. the 2 individuals being compared
# REMINDER: 'DATA' is full data set, 'FIT' based on a subset of the data of each site .-. range differs
# DATA = 1:43
# FIT_1 = 1:12

# Christoffer/Kyle
pair1 <- DATA[c(12,24)]
FIT_pair1 <- FIT_1[c(5,8)]
distance_pair1 <- distances(pair1, FIT_pair1) # distance measurement and time between the pair
distance_pair1$pair_ID <- "1"

# Christoffer/Elaine
pair2 <- DATA[c(12,15)]
FIT_pair2 <- FIT_1[c(5,6)]
distance_pair2 <- distances(pair2, FIT_pair2) # distance measurement and time between the pair
distance_pair2$pair_ID <- "2"

# Bumpus/Kyle
pair3 <- DATA[c(9,24)]
FIT_pair3 <- FIT_1[c(3,8)]
distance_pair3 <- distances(pair3, FIT_pair3) # distance measurement and time between the pair
distance_pair3$pair_ID <- "3"

# Elaine/Little Rick
pair4 <- DATA[c(15,27)]
FIT_pair4 <- FIT_1[c(6,9)]
distance_pair4 <- distances(pair4, FIT_pair4) # distance measurement and time between the pair
distance_pair4$pair_ID <- "4"

# Bumpus/Makao
pair5 <- DATA[c(9,29)]
FIT_pair5 <- FIT_1[c(3,10)]
distance_pair5 <- distances(pair5, FIT_pair5) # distance measurement and time between the pair
distance_pair5$pair_ID <- "5"

# Bumpus/Puji
pair6 <- DATA[c(9,36)]
FIT_pair6 <- FIT_1[c(3,11)]
distance_pair6 <- distances(pair6, FIT_pair6) # distance measurement and time between the pair
distance_pair6$pair_ID <- "6"

# Elaine/Rodolfo
pair7 <- DATA[c(15,38)]
FIT_pair7 <- FIT_1[c(6,12)]
distance_pair7 <- distances(pair7, FIT_pair7) # distance measurement and time between the pair
distance_pair7$pair_ID <- "7"

# Annie/Larry
pair8 <- DATA[c(2,25)] 
FIT_pair8 <- FIT_2[c(1,5)]
distance_pair8 <- distances(pair8, FIT_pair8) # distance measurement and time between the pair
distance_pair8$pair_ID <- "8"

# Larry/Reid
pair9 <- DATA[c(25,37)] 
FIT_pair9 <- FIT_2[c(5,9)]
distance_pair9 <- distances(pair9, FIT_pair9) # distance measurement and time between the pair
distance_pair9$pair_ID <- "9"

# Margaret/Thomas
pair10 <- DATA[c(30,42)] 
FIT_pair10 <- FIT_2[c(7,11)]
distance_pair10 <- distances(pair10, FIT_pair10) # distance measurement and time between the pair
distance_pair10$pair_ID <- "10"

# Reid/Thomas
pair11<- DATA[c(37,42)] 
FIT_pair11 <- FIT_2[c(9,11)]
distance_pair11 <- distances(pair11, FIT_pair11) # distance measurement and time between the pair
distance_pair11$pair_ID <- "11"

# Maria/Sheron
pair12 <- DATA[c(31,41)] 
FIT_pair12 <- FIT_2[c(8,10)]
distance_pair12 <- distances(pair12, FIT_pair12) # distance measurement and time between the pair
distance_pair12$pair_ID <- "12"

DISTANCE <- rbind(distance_pair1, distance_pair2, distance_pair3, distance_pair4, distance_pair5,
                  distance_pair6, distance_pair7, distance_pair8, distance_pair9, distance_pair10,
                  distance_pair11, distance_pair12)

#RDS of distances have been saved to "RDS/distance" folder
```



```{r}
ggplot(data = DATA_proximity, mapping = aes(x = sex_comparison, fill = sex_comparison)) + 
  geom_bar() +
  xlab("Sex") +
  #  ggtitle("HR Overlap pairwise comparison of sex") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), #removes horizontal gridlines
        panel.grid.minor = element_blank(), #removes vertical gridlines
        legend.position="none") + #removes legend
  scale_fill_manual(values = c("#A50026", "#9970AB", "#004488"),
                    labels = c("Female - Female", "Male - Female", "Male - Male"),
                    name = "")
```


## Correlative Movement

```{r}
#data carpentry
Bumpus <- DATA$Bumpus
Christoffer <- DATA$Christoffer
Elaine <- DATA$Elaine
Kyle <- DATA$Kyle
Little_rick <- DATA$`Little Rick`
Makao <- DATA$Makao
Puji <- DATA$Puji
Rodolfo <- DATA$Rodolfo
Annie <- DATA$Annie
Larry <- DATA$`Larry 267`
Margaret <- DATA$Margaret
Reid <- DATA$Reid
Thomas <- DATA$Thomas
Maria <- DATA$Maria
Sheron <- DATA$Sheron

#create a dataframe of an individuals GPS coordinates
Bumpus_GPS <- data.frame(timestamp = round_date(Bumpus$timestamp, "20 minutes"),
                          Bumpus.x = Bumpus$longitude,
                          Bumpus.y = Bumpus$latitude)
Christoffer_GPS <- data.frame(timestamp = round_date(Christoffer$timestamp, "20 minutes"),
                               Christoffer.x = Christoffer$longitude,
                               Christoffer.y = Christoffer$latitude)
Elaine_GPS <- data.frame(timestamp = round_date(Elaine$timestamp, "20 minutes"),
                          Elaine.x = Elaine$longitude,
                          Elaine.y = Elaine$latitude)
Kyle_GPS <- data.frame(timestamp = round_date(Kyle$timestamp, "20 minutes"),
                        Kyle.x = Kyle$longitude,
                        Kyle.y = Kyle$latitude)
Little_rick_GPS <- data.frame(timestamp = round_date(Little_rick$timestamp, "20 minutes"),
                               Little_rick.x = Little_rick$longitude,
                               Little_rick.y = Little_rick$latitude)
Makao_GPS <- data.frame(timestamp = round_date(Makao$timestamp, "20 minutes"),
                         Makao.x = Makao$longitude,
                         Makao.y = Makao$latitude)
Puji_GPS <- data.frame(timestamp = round_date(Puji$timestamp, "20 minutes"),
                        Puji.x = Puji$longitude,
                        Puji.y = Puji$latitude)
Rodolfo_GPS <- data.frame(timestamp = round_date(Rodolfo$timestamp, "20 minutes"),
                           Rodolfo.x = Rodolfo$longitude,
                           Rodolfo.y = Rodolfo$latitude)
Annie_GPS <- data.frame(timestamp = round_date(Annie$timestamp, "20 minutes"),
                         Annie.x = Annie$longitude,
                         Annie.y = Annie$latitude)
Larry_GPS <- data.frame(timestamp = round_date(Larry$timestamp, "20 minutes"),
                         Larry.x = Larry$longitude,
                         Larry.y = Larry$latitude)
Margaret_GPS <- data.frame(timestamp = round_date(Margaret$timestamp, "20 minutes"),
                            Margaret.x = Margaret$longitude,
                            Margaret.y = Margaret$latitude)
Reid_GPS <- data.frame(timestamp = round_date(Reid$timestamp, "20 minutes"),
                        Reid.x = Reid$longitude,
                        Reid.y = Reid$latitude)
Thomas_GPS <- data.frame(timestamp = round_date(Thomas$timestamp, "20 minutes"),
                          Thomas.x = Thomas$longitude,
                          Thomas.y = Thomas$latitude)
Maria_GPS <- data.frame(timestamp = round_date(Maria$timestamp, "20 minutes"),
                         Maria.x = Maria$longitude,
                         Maria.y = Maria$latitude)
Sheron_GPS <- data.frame(timestamp = round_date(Sheron$timestamp, "20 minutes"),
                          Sheron.x = Sheron$longitude,
                          Sheron.y = Sheron$latitude)

#combine the GPS coordinates of pairs together
pair1_GPS <- merge(Christoffer_GPS, Kyle_GPS)
pair1_GPS <- pair1_GPS[, c(1,2,4,3,5)]
pair1_GPS <- pair1_GPS[!duplicated(pair1_GPS$timestamp),]

pair2_GPS <- merge(Christoffer_GPS, Elaine_GPS)
pair2_GPS <- pair2_GPS[, c(1,2,4,3,5)]
pair2_GPS <- pair2_GPS[!duplicated(pair2_GPS$timestamp),]

pair3_GPS <- merge(Bumpus_GPS, Kyle_GPS)
pair3_GPS <- pair3_GPS[, c(1,2,4,3,5)]
pair3_GPS <- pair3_GPS[!duplicated(pair3_GPS$timestamp),]

pair4_GPS <- merge(Elaine_GPS, Little_rick_GPS)
pair4_GPS <- pair4_GPS[, c(1,2,4,3,5)]
pair4_GPS <- pair4_GPS[!duplicated(pair4_GPS$timestamp),]

pair5_GPS <- merge(Bumpus_GPS, Makao_GPS)
pair5_GPS <- pair5_GPS[, c(1,2,4,3,5)]
pair5_GPS <- pair5_GPS[!duplicated(pair5_GPS$timestamp),]

pair6_GPS <- merge(Bumpus_GPS, Puji_GPS)
pair6_GPS <- pair6_GPS[, c(1,2,4,3,5)]
pair6_GPS <- pair6_GPS[!duplicated(pair6_GPS$timestamp),]

pair7_GPS <- merge(Elaine_GPS, Rodolfo_GPS)
pair7_GPS <- pair7_GPS[, c(1,2,4,3,5)]
pair7_GPS <- pair7_GPS[!duplicated(pair7_GPS$timestamp),]

pair8_GPS <- merge(Annie_GPS, Larry_GPS)
pair8_GPS <- pair8_GPS[, c(1,2,4,3,5)]
pair8_GPS <- pair8_GPS[!duplicated(pair8_GPS$timestamp),]

pair9_GPS <- merge(Larry_GPS, Reid_GPS)
pair9_GPS <- pair9_GPS[, c(1,2,4,3,5)]
pair9_GPS <- pair9_GPS[!duplicated(pair9_GPS$timestamp),]

pair10_GPS <- merge(Margaret_GPS, Thomas_GPS)
pair10_GPS <- pair10_GPS[, c(1,2,4,3,5)]
pair10_GPS <- pair10_GPS[!duplicated(pair10_GPS$timestamp),]

pair11_GPS <- merge(Reid_GPS, Thomas_GPS)
pair11_GPS <- pair11_GPS[, c(1,2,4,3,5)]
pair11_GPS <- pair11_GPS[!duplicated(pair11_GPS$timestamp),]

pair12_GPS <- merge(Maria_GPS, Sheron_GPS)
pair12_GPS <- pair12_GPS[, c(1,2,4,3,5)]
pair12_GPS <- pair12_GPS[!duplicated(pair12_GPS$timestamp),]

```

### Correlative movement analysis

Determine if the movements are correlated in the pair. Calculate the correlative movement indices using the ```corrMove``` package.
Help files available for main functions in ```corrMove```.\
?as.corrData\
?findPrts\
?corrMove\
?plot.corrMove\

```{r}
#Create corrData object.
cdAnteater_pair1 <- as.corrData(pair1_GPS)
#Estimate the partition points
prtsAnteater_pair1 <- findPrts(cdAnteater_pair1, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair1 <- corrMove(cdAnteater_pair1, prtsAnteater_pair1)

#Create corrData object.
cdAnteater_pair2 <- as.corrData(pair2_GPS)
#Estimate the partition points
prtsAnteater_pair2 <- findPrts(cdAnteater_pair2, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair2 <- corrMove(cdAnteater_pair2, prtsAnteater_pair2)

#Create corrData object.
cdAnteater_pair3 <- as.corrData(pair3_GPS)
#Estimate the partition points
prtsAnteater_pair3 <- findPrts(cdAnteater_pair3, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair3 <- corrMove(cdAnteater_pair3, prtsAnteater_pair3)

#Create corrData object.
cdAnteater_pair4 <- as.corrData(pair4_GPS)
#Estimate the partition points
prtsAnteater_pair4 <- findPrts(cdAnteater_pair4, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair4 <- corrMove(cdAnteater_pair4, prtsAnteater_pair4)

#Create corrData object.
cdAnteater_pair5 <- as.corrData(pair5_GPS)
#Estimate the partition points
prtsAnteater_pair5 <- findPrts(cdAnteater_pair5, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair5 <- corrMove(cdAnteater_pair5, prtsAnteater_pair5)

#Create corrData object.
cdAnteater_pair6 <- as.corrData(pair6_GPS)
#Estimate the partition points
prtsAnteater_pair6 <- findPrts(cdAnteater_pair6, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair6 <- corrMove(cdAnteater_pair6, prtsAnteater_pair6)

#Create corrData object.
cdAnteater_pair7 <- as.corrData(pair7_GPS)
#Estimate the partition points 
prtsAnteater_pair7 <- findPrts(cdAnteater_pair7, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair7 <- corrMove(cdAnteater_pair7, prtsAnteater_pair7)

#Create corrData object.
cdAnteater_pair8 <- as.corrData(pair8_GPS)
#Estimate the partition points
prtsAnteater_pair8 <- findPrts(cdAnteater_pair8, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair8 <- corrMove(cdAnteater_pair8, prtsAnteater_pair8)

#Create corrData object.
cdAnteater_pair9 <- as.corrData(pair9_GPS)
#Estimate the partition points
prtsAnteater_pair9 <- findPrts(cdAnteater_pair9, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair9 <- corrMove(cdAnteater_pair9, prtsAnteater_pair9)

#Create corrData object.
cdAnteater_pair10 <- as.corrData(pair10_GPS)
#Estimate the partition points
prtsAnteater_pair10 <- findPrts(cdAnteater_pair10, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair10 <- corrMove(cdAnteater_pair10, prtsAnteater_pair10)

#Create corrData object.
cdAnteater_pair11 <- as.corrData(pair11_GPS)
#Estimate the partition points
prtsAnteater_pair11 <- findPrts(cdAnteater_pair11, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair11 <- corrMove(cdAnteater_pair11, prtsAnteater_pair11)

#Create corrData object.
cdAnteater_pair12 <- as.corrData(pair12_GPS)
#Estimate the partition points
prtsAnteater_pair12 <- findPrts(cdAnteater_pair12, W=5, IC = 2)
#Get the MCI estimates and selected model conditional on the data and partition points
cmAnteater_pair12 <- corrMove(cdAnteater_pair12, prtsAnteater_pair12)

#RDS of correlative movement have been saved to "RDS/correlative movement" folder
```


